---
title: "Tutorial 7 Lesson 1"
author: "Lucia Vilallonga"
date: "9/17/2025"
output: html_document
---

echo=FALSE ?
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


# Exercise 0: Load data
Load the "icecream.csv" data file into a DataFrame object.
```{r}
```

Solution:
```{r}
icecream <- read.csv("icecream.csv")
```


# Exercise 1: Variable Exploration
Plot the relationships between the following variables: 
(a) ice cream consumption vs outside temperature
```{r}
```

(b) ice cream consumption vs ice cream price, colored by outside temperature
```{r}
```

(c) ice cream consumption vs consumer income, colored by outside temperature
```{r}
```

Solution:
```{r}
ggplot(data=icecream, aes(y=cons, x=temp)) +
  geom_point()

ggplot(data=icecream, aes(y=cons, x=price, col=temp)) +
  geom_point()

ggplot(data=icecream, aes(y=cons, x=income, col=temp)) +
  geom_point()
```


# Exercise 2: How would you describe the relationship between each of the variables above? 
You can use terms like 'positive/negative/no relationship,' 'strong/weak/no correlation,' 'linear/nonlinear,' and anything else you think describes the data.

Solution: ***TODO: write***
(a) 
(b)
(c)


# Exercise 3: Writing the model
Write by hand a linear model for this problem, with the variables
explored above. Use beta's as needed for the model parameters, and state 
clearly which variables you're using and how you're naming them.

Solution: ***TODO: write***


# Exercise 4: Implement the model in R with lm()
```{r}
```

Solution:
```{r}
m1 <- lm(data=icecream, cons ~ income + price + temp)
summary(m1)
```


# Exercise 5: ***TODO: write, like Exercise 4 in IMS supp***


# Exercise 6: Add interaction terms
Implement a new model with interaction between income and price.
```{r}
```

Solution:
```{r}
m2 <- lm(data=icecream, cons ~ income + price + temp + income:price)
summary(m2)
```


# Exercise 7: Write the interactions model above

Solution: ***TODO: write***


# Exercise 8: ***TODO: write, like Exercise 6 in IMS supp***


