---
title: "Advanced Inference: 1 - Interactions"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
#devtools::install_github("rundel/learnrhash")
#devtools::install_git("git@github.com:rundel/learnrhash.git")

library(learnr)
library(tidyverse)
library(openintro)
library(grid)
library(png)
library(ggplot2)
#library(emo)

knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center",
                      fig.height = 3,
                      fig.width = 5,
                      message = FALSE,
                      warning = FALSE)

tutorial_options(exercise.eval = FALSE)

# Hash generation helpers
# Should ideally be loaded from the imstutorials package when it exists
is_server_context <- function(.envir) {
  # We are in the server context if there are the follow:
  # * input - input reactive values
  # * output - shiny output
  # * session - shiny session
  #
  # Check context by examining the class of each of these.
  # If any is missing then it will be a NULL which will fail.
  
  inherits(.envir$input, "reactivevalues") &
    inherits(.envir$output, "shinyoutput") &
    inherits(.envir$session, "ShinySession")
}

check_server_context <- function(.envir) {
  if (!is_server_context(.envir)) {
    calling_func <- deparse(sys.calls()[[sys.nframe() - 1]])
    err <- paste0("Function `", calling_func, "`", " must be called from an Rmd chunk where `context = \"server\"`")
    stop(err, call. = FALSE)
  }
}
encoder_logic <- function(strip_output = FALSE) {
  p <- parent.frame()
  check_server_context(p)
  # Make this var available within the local context below
  assign("strip_output", strip_output, envir = p)
  # Evaluate in parent frame to get input, output, and session
  local(
    {
      encoded_txt <- shiny::eventReactive(
        input$hash_generate,
        {
          # shiny::getDefaultReactiveDomain()$userData$tutorial_state
          state <- learnr:::get_tutorial_state()
          shiny::validate(shiny::need(length(state) > 0, "No progress yet."))
          shiny::validate(shiny::need(nchar(input$name) > 0, "No name entered."))
          shiny::validate(shiny::need(nchar(input$studentID) > 0, "Please enter your student ID"))
          user_state <- purrr::map_dfr(state, identity, .id = "label")
          user_state <- dplyr::group_by(user_state, label, type, correct)
          user_state <- dplyr::summarize(
            user_state,
            answer = list(answer),
            timestamp = dplyr::first(timestamp),
            .groups = "drop"
          )
          user_state <- dplyr::relocate(user_state, correct, .before = timestamp)
          user_info <- tibble(
            label = c("student_name", "student_id"),
            type = "identifier",
            answer = as.list(c(input$name, input$studentID)),
            timestamp = format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z", tz = "UTC")
          )
          learnrhash::encode_obj(bind_rows(user_info, user_state))
        }
      )
      output$hash_output <- shiny::renderText(encoded_txt())
    },
    envir = p
  )
}

hash_encoder_ui <- {
  shiny::div("If you have completed this tutorial and are happy with all of your", "solutions, please enter your identifying information, then click the button below to generate your hash", textInput("name", "What's your name?"), textInput("studentID", "What is your student ID?"), renderText({
    input$caption
  }), )
}
```

## Welcome

Previously, you've learned how to perform multiple regression for inference and prediction. We will build on those skills in this tutorial, and in particular learn how to adapt models where two or more variables interact with each other and the outcome of interest.

As a simple example, suppose we want to predict how much food cats will eat in a week. We have a sample of 16 pet cats, and have recorded their ages, weights, and body lengths, as well as how many grams of food they eat. Many of these variables are inter-correlated: kittens eat less than adult cats, and they are also smaller by weight and length. Similarly, many larger cats by length also weigh more (and eat more) than their shorter peers. In a situation like this, it's difficult to tease apart the relationships between the covariates and the outcome. We will learn how to address this below. 


### Example: demographic data

An academic advisor wants to study the career outcomes of former students under her mentorship. She collects the following data on 16 recent graduates: 

- `income` - annual income, in tens of thousands of USD
- `experience` - years of work experience
- `gender` - factor indicating gender 

Here's what a scatterplot of the data looks like:

```{r, echo=FALSE}
inc <- c(43, 48, 52, 70, 61, 83, 96, 100,
         36, 40, 39, 44, 46, 49, 50, 53) 
exp <- c(1, 2, 4, 6, 5, 3, 8, 10,
         1, 3, 2, 5, 8, 6, 7, 10)
gen <- c("male", "male", "male", "male", "male", "male", "male", "male",
         "female", "female", "female", "female", "female", "female", "female", 
         "female")

data <- data.frame(inc, exp, gen)
colnames(data) <- c("income", "experience", "gender")

ggplot(data=data, aes(y=income, x=experience, col=gender)) +
  geom_point()
```


## Linear model for demographic data

Let's fit a linear model to these data, using all the covariates we have: 

```{r, include=TRUE}
m1 <- lm(data=data, income ~ .)
summary(m1)

# TODO: add plot and discuss deficiency
```

The plots below show how well the model without an interaction term fits the data:

```{r, include=TRUE}
ggplot(data=data, aes(x=experience, y=income)) +
  geom_point(aes(col=gender)) +
  geom_smooth(method='lm', se=FALSE) +
  ggtitle('Linear model without interaction term')

data$residuals <- resid(m1)
data$fitted <- fitted(m1)

ggplot(data=data, aes(x=fitted, y=residuals)) +
  geom_point(aes(col=gender)) +
  geom_hline(yintercept=0) +
  ggtitle('Residuals vs fitted values for model without interaction term')

ggplot(data=data, aes(sample=residuals)) +
  stat_qq() + 
  stat_qq_line() +
  ggtitle('QQ plot for model without interaction term') +
  ylab("residuals") +
  xlab("quantiles")
```

We can see from the plots above that key linearity assumptions are violated and the model doesn't fit very well:

1. The variance of the residuals is non-constant (especially across the gender groups). More specifically, the residuals are larger for the male income values; and,
2. The quantiles in the QQ plot deviate from the diagonal, indicating that the residuals are not normally distributed.

This indicates that we might have to try something else!


## Adding interactions to our model

In R, the syntax for adding an interaction term within the linear model `lm()` is `lm(y ~ x1 + x2 + x1:x2)`, where $y$ is the dependent variable and $x_1, x_2$ are two covariates. Let's try it for our demographic data:

```{r, include=TRUE}
m2 <- lm(data=data, income ~ experience + gender + experience:gender)
summary(m2)
```

And now we have the following plots for `m2`:

```{r, include=TRUE}
data$residuals <- resid(m2)
data$fitted <- fitted(m2)

ggplot(data=data, aes(x=fitted, y=residuals)) +
  geom_point(aes(col=gender)) +
  geom_hline(yintercept=0) +
  ggtitle('Residuals vs fitted values for model with interaction term')

ggplot(data=data, aes(sample=residuals)) +
  stat_qq() + 
  stat_qq_line() +
  ggtitle('QQ Plot for model with interaction term')  +
  ylab("residuals") +
  xlab("quantiles")
```

This model is also not perfect (check: why?), but let's see how the two compare. 

### Comparing model fit

```{r, echo=FALSE}
tab <- matrix(c(0.733, 0.692, 11.1, 0.833, 0.791, 9.1), ncol=3, byrow=TRUE)
colnames(tab) <- c('Multiple R-squared', 'Adjusted R-squared', 'Residual SE')
rownames(tab) <- c('m1', 'm2')
tab <- as.table(tab)
print(tab)
```

We can see that the model that includes an interaction term performs better according to adjusted $R^2$, which takes into account the complexity in the model that is added when we add a term.  


## Your turn!

### Data exploration

Now we'll turn to a more interesting problem that does not have a categorical variable, but where our model can still benefit from including interactions. In this analysis, we are interested in determining what factors influence ice cream consumption from an ice cream truck with a route that travels across different neighborhoods. The variables are:

- `cons` - ice cream consumption
- `temp` - outdoor temperature
- `income` - average neighborhood income
- `price` - ice cream price

To start, load the `icecream.csv` data file into a DataFrame and create scatterplots to examine the relationships between the varables. 

```{r ex1, exercise = TRUE}
# Load data
icecream <- _______("data/icecream.csv")

# View some of the variables
head(icecream)

# Plot ice cream consumption vs outside temperature
ggplot(data=icecream, aes(y=_______, x=_______)) +
  geom_point()

# Plot ice cream consumption vs ice cream price, colored by outside temperature
ggplot(data=icecream, aes(y=_______, x=_______, col=temp)) +
  geom_point()

# Plot ice cream consumption vs consumer income, colored by price
ggplot(data=icecream, aes(y=_______, x=_______, col=_______)) +
  geom_point()
```

<details>
<summary>**Hint**</summary>
Try `read.csv()` for loading the csv file.
</details>

<!-- ```{r ex1-solution} -->
<!-- # Solution -->
<!-- icecream <- read.csv("data/icecream.csv") -->

<!-- ggplot(data=icecream, aes(y=cons, x=temp)) + -->
<!--   geom_point() -->

<!-- ggplot(data=icecream, aes(y=cons, x=price, col=temp)) + -->
<!--   geom_point() -->

<!-- ggplot(data=icecream, aes(y=cons, x=income, col=temp)) + -->
<!--   geom_point() -->
<!-- ``` -->

```{r mc1}
question("What is the relationship between ice cream consumption and temperature?",
         answer("Positive and somewhat linear", correct=TRUE),
         answer("Positive and strongly linear", message="Not quite. The relationship is positive, but check its strength again."),
         answer("Negative and non-linear", message="Incorrect. Recall, what does it look like for two variables to be positive or negatively correlated?"),
         answer("There is no correlation", message="Incorrect. There is clearly some correlation, check again to see what the correct description would be."),
         allow_retry = TRUE)
```

### Modeling icecream consumption

We can see that ice cream consumption and temperature have a positive and somewhat linear relationship; that is, when temperature increases, we observe that consumption is also likely to increase. But, we have two other variables in the dataset: price and income. Could these two variables be inter-related?

First, let's write a model for the data: if $y$ is consumption, then with the variables we explored above the model can be expressed as

$y = \beta_0 + \beta_{\rm temp}x_{\rm temp} + \beta_{\rm price}x_{\rm price} + \beta_{\rm income}x_{\rm income} + residuals$

Now, fit this model below:

```{r ex2, exercise = TRUE}
m1 <- ____

summary(m1)
```

<!-- ```{r ex2-solution} -->
<!-- # Add the independent variables (covariates) after the ~ and separated by + -->
<!-- m1 <- lm(data=icecream, cons ~ temp + price + income) -->

<!-- summary(m1) -->
<!-- ``` -->


### Interactions between non-categorical variables

Now add the interaction term corresponding to the written model below: 

$y = \beta_0 + \beta_{\rm temp}x_{\rm temp} + \beta_{\rm price}x_{\rm price} + \beta_{\rm income}x_{\rm income} + \beta_{\rm price:income}x_{\rm price}x_{\rm income} + residuals$

Note that all of these are continuous variables, not categorical. 

```{r ex3, exercise = TRUE}
m2 <- ____

summary(m2)
```

<!-- ```{r ex3-solution} -->
<!-- # Add the independent variables (covariates) after the ~ and separated by + -->
<!-- m2 <- lm(data=icecream, cons ~ temp + price + income + price:income) -->

<!-- summary(m2) -->
<!-- ``` -->


### Assessing model fit

```{r, echo=FALSE}
tab <- matrix(c(0.719, 0.687, 0.037, 0.759, 0.721, 0.035), ncol=3, byrow=TRUE)
colnames(tab) <- c('Multiple R-squared', 'Adjusted R-squared', 'Residual SE')
rownames(tab) <- c('m1', 'm2')
tab <- as.table(tab)
print(tab)
```

```{r mc2}
question("Given the measures of model fit above, which model do you think fits the icecream consumption data better?",
         answer("m1, linear model with no interactions", message="Incorrect. Try checking the diagnostics again, especially the adjusted R^2."),
         answer("m2, linear model with interaction between price and income", correct=TRUE),
         allow_retry = TRUE)
```

## Inference for interaction terms

Now that we've added an interaction term to the model, we may ask whether it is meaningful or not. That is, we want to conduct the following (two-sided) hypothesis test:

> H0: $\beta_{\rm price:income} = 0$

> H1: $\beta_{\rm price:income} \neq 0$

###

Recall the model summary:

```{r, include=TRUE}
summary(m2)$coefficients
```

The final column, `Pr(>|t|)`, gives the probability of the interaction effect being at least as extreme as the one observed (i.e., $\mathbb{P}(\beta_{\rm price:income} \geq |4.26|)$) _assuming H0 is true_. In other words, this is exactly the p-value. 

```{r mc3}
question("Given the p-value of 0.02, is the interaction effect meaningful at confidence level alpha=0.05?",
         answer("Yes, we can reject H0.", correct=TRUE, message="Right, but with a caveat. The p-value is only slighly less than 0.05, which suggests that the interaction term *might be* significant in this context. To be more sure, we would want to see a much smaller p-value."),
         answer("No, we don't have enough evidence to reject H0.", message="Incorrect - check again. When would you reject H0?"),
         allow_retry = TRUE)
```

## Submit

```{r, echo=FALSE, context="server"}
encoder_logic()
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(ui_before = hash_encoder_ui)
```
