---
title: "Advanced Inference: 2 - Nonlinear Regression"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
#devtools::install_github("rundel/learnrhash")
#devtools::install_git("git@github.com:rundel/learnrhash.git")

library(learnr)
library(tidyverse)
library(openintro)
library(grid)
library(png)
#library(emo)

knitr::opts_chunk$set(echo = FALSE,
                      fig.align = "center",
                      fig.height = 3,
                      fig.width = 5,
                      message = FALSE,
                      warning = FALSE)

tutorial_options(exercise.eval = FALSE)

# Hash generation helpers
# Should ideally be loaded from the imstutorials package when it exists
is_server_context <- function(.envir) {
  # We are in the server context if there are the follow:
  # * input - input reactive values
  # * output - shiny output
  # * session - shiny session
  #
  # Check context by examining the class of each of these.
  # If any is missing then it will be a NULL which will fail.
  
  inherits(.envir$input, "reactivevalues") &
    inherits(.envir$output, "shinyoutput") &
    inherits(.envir$session, "ShinySession")
}

check_server_context <- function(.envir) {
  if (!is_server_context(.envir)) {
    calling_func <- deparse(sys.calls()[[sys.nframe() - 1]])
    err <- paste0("Function `", calling_func, "`", " must be called from an Rmd chunk where `context = \"server\"`")
    stop(err, call. = FALSE)
  }
}
encoder_logic <- function(strip_output = FALSE) {
  p <- parent.frame()
  check_server_context(p)
  # Make this var available within the local context below
  assign("strip_output", strip_output, envir = p)
  # Evaluate in parent frame to get input, output, and session
  local(
    {
      encoded_txt <- shiny::eventReactive(
        input$hash_generate,
        {
          # shiny::getDefaultReactiveDomain()$userData$tutorial_state
          state <- learnr:::get_tutorial_state()
          shiny::validate(shiny::need(length(state) > 0, "No progress yet."))
          shiny::validate(shiny::need(nchar(input$name) > 0, "No name entered."))
          shiny::validate(shiny::need(nchar(input$studentID) > 0, "Please enter your student ID"))
          user_state <- purrr::map_dfr(state, identity, .id = "label")
          user_state <- dplyr::group_by(user_state, label, type, correct)
          user_state <- dplyr::summarize(
            user_state,
            answer = list(answer),
            timestamp = dplyr::first(timestamp),
            .groups = "drop"
          )
          user_state <- dplyr::relocate(user_state, correct, .before = timestamp)
          user_info <- tibble(
            label = c("student_name", "student_id"),
            type = "identifier",
            answer = as.list(c(input$name, input$studentID)),
            timestamp = format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z", tz = "UTC")
          )
          learnrhash::encode_obj(bind_rows(user_info, user_state))
        }
      )
      output$hash_output <- shiny::renderText(encoded_txt())
    },
    envir = p
  )
}

hash_encoder_ui <- {
  shiny::div("If you have completed this tutorial and are happy with all of your", "solutions, please enter your identifying information, then click the button below to generate your hash", textInput("name", "What's your name?"), textInput("studentID", "What is your student ID?"), renderText({
    input$caption
  }), )
}
```

## Welcome

In this lesson, we'll learn how to model data that may not be linear. See the data table below:

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)  
library(ggplot2)

# Simulate some data 
x <- runif(n=100, min=0, max=10)
epsilon <- rnorm(n=100, mean=0, sd=1)
y <- 2*x**2 + epsilon*((1+x**2))

data <- data.frame(x,y)
head(data)
```


### Visualize data

Let's examine the data. First, generate a scatterplot of x and y.
```{r}
ggplot(data=data, aes(x=x, y=y)) +
  geom_point()
```

```{r mc1, echo=FALSE}
question("Is the relationship between x and y linear?",
         answer("Yes"),
         answer("No", correct=TRUE),
         answer("Not sure"))
```

Now let's look at the histograms of x and y, below. Notice that x looks more or less uniformly distributed, but y is highly skewed. This is another clue that the relationship between the variables is not linear.

```{r}
ggplot(data=data, aes(x=x)) +
  geom_histogram()

ggplot(data=data, aes(x=y)) +
  geom_histogram()
```


### Adding a transformation

**Exercise**

Compute y_ = sqrt(y) and add it to the dataframe. Then, plot the relationship between x and y_.
```{r ex2, exercise=TRUE}
data$y_ <- _____

ggplot(data=data, aes(x=_____, y=_____)) +
  geom_point()
```

<!-- ```{r ex2-solution} -->
<!-- data$y_ = sqrt(data$y) -->

<!-- ggplot(data=data, aes(x=x, y=y_)) + -->
<!--   geom_point() -->
<!-- ``` -->

Does the relationship between x and y_ look linear?
```{r mc2, echo=FALSE}
question("Is the relationship between x and y_ linear?",
         answer("Yes", correct=TRUE),
         answer("No"),
         answer("Not sure"))
```


### Modeling the transformed data

**Exercise**

Now, fit a linear model for x and y_. 

```{r ex3, exercise=TRUE}
model <- lm(______)
summary(model)
```

<!-- ```{r ex3-solution} -->
<!-- model <- lm(data=data, y_ ~ x) -->
<!-- summary(model) -->
<!-- ``` -->


### Back-transformation

Now we need to back-transform the data in order to plot the best-fit line on
the original (non-transformed) domain of y. 

$y* = 0.557 + 1.268x + residuals$

$\sqrt{y}= 0.557 + 1.268x + residuals$

$y = (0.557 + 1.268x + residuals)^2$

The model on the last line is shown below.

```{r}
ggplot(data=data, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method="lm", formula=y~poly(x,2), se=FALSE)
```

Note: after transforming the data, you will need to take care when interpreting the coefficients of the model. As we showed, we cannot describe changes in $y$ as linear with respect to changes in $x$. Instead, in this specific case the change in $y$ is *quadratic* with changes in $x$. 


## Your turn!

Now try repeating the same analysis on MLB salary data, specifically for just the 1986 and 1987 seasons. Here's a preview of what the dataset, called `Hitters` ($N=263$),  looks like:

```{r, echo=FALSE}
library(ISLR2)
library(dplyr)

data(Hitters)

Hitters <- Hitters %>% 
  select(Salary, RBI) %>%
  filter(!is.na(Salary)) %>%
  filter(Salary > 0)

head(Hitters)
```

We are interested in modeling these players' salaries vs RBI. RBI, coarsely speaking, measures how many points the team scores as a result of a player's at-bat. The units of the `Salary` column are in thousands of dollars.

**Exercise**

Create a scatterplot of players' salaries vs RBI. Check the linearity of the data with the scatterplot.

```{r ex4, exercise=TRUE}
```

<!-- ```{r ex4-solution} -->
<!-- ggplot(data=Hitters, aes(x=RBI, y=Salary)) + -->
<!--   geom_point() -->
<!-- ``` -->

###

### A different transformation

So the Salaries are not linear, and you'll have to transform them again. Take another look at the scatterplot above.

```{r mc3, echo=FALSE}
question("What relationship does MLB players' salary and RBI seem to have?",
         answer("Linear"),
         answer("Exponential", correct=TRUE),
         answer("Cubic"),
         answer("Not sure"))
```


### Modeling exponential data

**Exercise**

Perform a log transformation on the Salary variable, and add it to the dataframe. Then use the transformed variable in a linear model as you did with the quadratic example above.

```{r ex5, exercise=TRUE}
# Perform a log transformation on the salary variable
Hitters$logSalary <- _____

# Create a linear model of log-salary vs RBI
m2 <- ____
summary(m2)
```

<!-- ```{r ex5-solution} -->
<!-- Hitters$logSalary <- log(Hitters$Salary) -->

<!-- m2 <- lm(data=Hitters, logSalary ~ RBI) -->
<!-- summary(m2) -->
<!-- ``` -->

```{r mc4, echo=FALSE}
question("What inverse operation will you need to perform on the model for the correct back-transformation?",
         answer("(...)^2"),
         answer("exp(...) or e^(...)", correct=TRUE),
         answer("sqrt(...)"),
         answer("1/(...)"))
```


### Back transformation

After back-transforming, your model will be:

$y = e^{\beta_0 + \beta_1x_{RBI} + residuals}$

### Assumptions and model diagnostics

Ok, so now `logSalary` appears linear with `RBI`, at least visually. But, we still need to check our assumptions and diagnose the model fit. The only thing that changes now is that we're doing this with respect to the transformed variable.

**Exercise**

Using your `m2` object above, check how well your transformed model fits the data:

1. Create a plot of the residuals vs the fitted values, with a y-intercept at 0
2. Create a qq-plot of the residuals

```{r ex6, exercise=TRUE}
# Extract the residuals and fitted values from the m2 object
Hitters$residuals <- ____
Hitters$fitted <- ____

# Plot residuals vs fitted values, with a yintercept at 0
ggplot(data=Hitters, aes(x=____, y=____)) +
  geom_hline(yintercept=____) +
  ggtitle('Residuals vs fitted values for m2')

# QQ-plot of residuals
ggplot(data=Hitters, aes(sample=residuals)) +
  stat_qq() + 
  stat_qq_line() +
  ggtitle('QQ Plot for m2')  +
  ylab("residuals") +
  xlab("quantiles")
```

<!-- ```{r ex6-solution} -->
<!-- # Extract the residuals and fitted values from the m2 object -->
<!-- Hitters$residuals <- resid(m2) -->
<!-- Hitters$fitted <- fitted(m2) -->

<!-- # Plot residuals vs fitted values, with a yintercept at 0 -->
<!-- ggplot(data=Hitters, aes(x=fitted, y=residuals)) + -->
<!--   geom_point() + -->
<!--   geom_hline(yintercept=0) + -->
<!--   ggtitle('Residuals vs fitted values for m2') -->

<!-- # QQ-plot of residuals -->
<!-- ggplot(data=Hitters, aes(sample=residuals)) + -->
<!--   stat_qq() +  -->
<!--   stat_qq_line() + -->
<!--   ggtitle('QQ Plot for m2')  + -->
<!--   ylab("residuals") + -->
<!--   xlab("quantiles") -->
<!-- ``` -->


```{r mc6, echo=FALSE}
question("Given the plots above, what assumptions (if any) are violated for modeling logSalary vs RBI using linear regression?",
         answer("No assumptions are violated, everything looks good.", message="Try checking the plots again. What would you want to see in a graph of the residuals vs fitted values, and in a QQ plot?"),
         answer("As fitted values increase, the values of the residuals decrease, indicating that they have a non-constant variance.", message="Yes, this is true. Does anything appear off about the QQ plot?"),
         answer("The quantiles of the residuals do not follow the diagonal line, which indicates that they are not normally distributed.", message="Yes, this is true. Does anything appear off about the plot of residuals vs fitted values?"),
         answer("The residuals seem to have a non-constant variance and they are not normally distributed.", correct=TRUE),
         allow_retry = TRUE)
```


### Interpretation

Now on a log scale, `logSalary` appears linear with RBI:

```{r, include=FALSE}
Hitters$logSalary <- log(Hitters$Salary)
```

```{r}
ggplot(data=Hitters, aes(x=RBI, y=logSalary)) +
  geom_point() +
  geom_smooth(method="lm", formula=(y~x), se=FALSE, color="blue")
```

Note the difference in the scales on the y-axis compared to the first scatterplot you made, because now we're on a log scale. 

```{r mc5, echo=FALSE}
question("Using the model summary below, how would you interpret the coefficient for RBI?",
         answer("Every unit increase in RBI increases the players' salary by 0.01529 thousands of dollars (or $15.259).", message="This is an additive interpretation, which would be true if we were modeling Salary vs RBI - but recall that we used a log transformation. Try looking at the back-transformed model again."), # (0.015259*1000 =) $15.26.
         answer("Every unit increase in RBI increases the players' salary by a factor of exp(0.015259) = 1.015.", correct=TRUE), # (exp(0.015259)*1000 = 1.015259*1000 =) $1015.38
         answer("Every unit increase in RBI increases the players' salary by a factor of 0.015259.", message="Careful - the effect is multiplicative but be sure to double-check the scale."),
         answer("Every unit increase in RBI increases the players' salary by a factor of 1.5259%.", message="Careful - the effect is multiplicative but recall that we need to use exp() to back-transform the model. Try computing exp(B1(x+1))/exp(B1x) and see what you get."),
         allow_retry = TRUE)
```

```{r, include=FALSE}
m2 <- lm(data=Hitters, logSalary ~ RBI)
summary(m2)
```




## Submit

```{r, echo=FALSE, context="server"}
encoder_logic()
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(ui_before = hash_encoder_ui)
```
